---
title: "My Fancy Presentation Made Using R Markdown"
author: "James Ibbetson"
date: "8/29/2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

For a [cheatsheet](www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf).

For a [reference guide](www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf).

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## My slide showing a few formatting options

in plain text

*in italicized text*

**in bold text**

`in code format, e.g. for (i %in% 1:10)`

## My slide with headings, bullet lists, and number lists
### A top level subheading
#### A second level subheading
- Bullet one
- Bullet two
- Bullet three

1. Item a
2. Item b
3. Item c

## My slide with R code
First without showing the code...
```{r}
head(mtcars, 3)
```

...and now with the code
```{r, echo = TRUE}
head(mtcars, 3)
```
## My slide with more R code options
Now with no ## marks for each line
```{r, comment = "", background = "blue" }
head(mtcars, 3)
```

Just the code without evaluation
```{r, echo = TRUE, eval = FALSE}
head(mtcars, 3)
```

***
My slide without a title

## My slide with plots

```{r pressure}
par(mfrow=c(1,2))
    plot(pressure)
    plot(pressure, col='red')
```

