---
title: "Playing with Leaflet"
author: "James Ibbetson"
date: "8/29/2017"
always_allow_html: yes
---

### *Disclaimer*
*The following notes draw heavily from the class lectures and Rmd file covering Leaflet. For the original material and appropriate credit please see the class [Lectures and source code](https://github.com/DataScienceSpecialization/Developing_Data_Products).*

## Introduction

Leaflet is one of the most popular Javascript libraries for
creating interactive maps. The leaflet R package allows you
to create your own leaflet maps without needing to know any
Javascript!

```{r, eval=FALSE}
install.packages("leaflet")
library(leaflet)
```

```{r, echo=FALSE}
library(leaflet)
```

## The basics

Getting started with leaflet is easy. The `leaflet()`
function creates a map widget that you can store in a
variable so that you can modify the map later on. 
```{r, comment = "" }
my_map <- leaflet()
class(my_map)
```

Right now `my_map` is empty. You can add features to the map using 
the pipe operator (`%>%`) just like in dplyr. The `addTiles()` function 
adds mapping data from [Open Street Map](http://www.openstreetmap.org/). 
Add markers to your map using the `addMarkers()` function by 
specifying the longitude and latitude. Add the popup text for when you 
click on the marker with the `popup` argument in `addMarkers()`.

```{r, eval=FALSE}
my_map <- leaflet() %>% 
    addTiles() %>%
    addMarkers(lat=53.430823, lng=-2.960835, popup="This is Anfield")
my_map
```

```{r, echo=FALSE}
my_map <- leaflet() %>% 
    addTiles() %>%
    addMarkers(lat=53.430823, lng=-2.960835, popup="This is Anfield")
my_map
```

## Working with dataframes
The leaflet package in R is designed to work efficiently with dataframes as inputs for `leaflet()` and its related functions.
Data for multiple markers and their associated information can be passed through 
directly as long as there are columns for latitude and longitude. You can even 
add hyperlinks to the marker popups. Plus many more options...

Here's an example using a simple dataframe with five entries:
```{r, eval=FALSE}
uk_fun <- data.frame(name = c("Churchill War Rooms", "High Ropes Oxford", "Bletchley Park",
                              "Stonehenge", "Berkhamsted Castle"),
                      lat = c(51.502264, 51.653377, 51.997660, 51.178828, 51.763617),
                      lng = c(-0.129039, -1.232407, -0.740686, -1.826247, -0.559281),
                    sites = c(
  "<a href='http://www.iwm.org.uk/visits/churchill-war-rooms'>Churchill War Rooms</a>",
  "<a href='http://www.highropesoxford.com/'>High Ropes Oxford</a>",
  "<a href='https://www.bletchleypark.org.uk'>Bletchley Park</a>",
  "<a href='http://www.english-heritage.org.uk/visit/places/stonehenge/'>Stonehenge</a>",
  "Berkhamsted Castle"))
uk_fun %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = uk_fun$sites) %>%
  setView(lng=-1.2, lat=51.45, zoom = 8)  # specifies map center & zoom
```

```{r, echo=FALSE}
uk_fun <- data.frame(name = c("Churchill War Rooms", "High Ropes Oxford", "Bletchley Park",
                              "Stonehenge", "Berkhamsted Castle"),
                      lat = c(51.502264, 51.653377, 51.997660, 51.178828, 51.763617),
                      lng = c(-0.129039, -1.232407, -0.740686, -1.826247, -0.559281),
                    sites = c(
  "<a href='http://www.iwm.org.uk/visits/churchill-war-rooms'>Churchill War Rooms</a>",
  "<a href='http://www.highropesoxford.com/'>High Ropes Oxford</a>",
  "<a href='https://www.bletchleypark.org.uk'>Bletchley Park</a>",
  "<a href='http://www.english-heritage.org.uk/visit/places/stonehenge/'>Stonehenge</a>",
  "Berkhamsted Castle"))
uk_fun %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = uk_fun$sites) %>%
  setView(lng=-1.2, lat=51.45, zoom = 8)  # specifies map center & zoom
```

## Suppressing errors and warnings
The example above works but notice the annoying `## Assuming 'lng' and 'lat' are longitude and latitude, respectively` message in the output. To suppress this, use the `message` option `{r, eval=FALSE, warning=FALSE, message=FALSE}` in the chunk header.

## More popup details
The following is code I reviewed from a [classmate's leaflet assignment](http://rpubs.com/DanielD/302176). Note also the insertion of linebreaks in the html output, which can be achieved one of two ways: either type \<br>\ or add two spaces at the end of a line.

*Create the fields that will be used to provide the following details for the map:  
-The latitude and longitude of each school location  
-The details that will appear in the popup when a school location is clicked  
-The colour of the circle that will appear around each school*  

```{r, eval=FALSE}
latLng <- data.frame(lat=schoolEnrolments$Y, lng=schoolEnrolments$X)

schoolPopup <- paste0("School Name: ", schoolEnrolments$School_Name.x, "<br>",
                      "Town: ", schoolEnrolments$Address_Town, "<br>",
                      "Type: ", schoolEnrolments$School_Type.x, "<br>",
                      "Sector: ", schoolEnrolments$Education_Sector.x, "<br>",
                      "Total Enrolments (FTE): ", schoolEnrolments$Grand.Total)

schoolColour <- ifelse(schoolEnrolments$Grand.Total < 500, "green",
                       ifelse(schoolEnrolments$Grand.Total < 1000, "blue",
                              ifelse(schoolEnrolments$Grand.Total < 2000, "orange", "red")))
```

*Show the Interactive Map using the details defined and set the following:<br>
- clustering of schools<br>
- the size of the coloured circle around each school<br>
- a legend of the enrolment numbers used in setting the colours of the circles around each school*<br>
```{r, eval=FALSE}
latLng %>%
    leaflet() %>%
    addTiles() %>%
    addMarkers(clusterOptions = markerClusterOptions(), icon = schoolIcon, popup = schoolPopup) %>%
    addCircles(weight = 2, radius = sqrt(schoolEnrolments$Grand.Total) * 2, color = schoolColour) %>%
    addLegend(title = "Total Enrolments (FTE)",
              labels = c("< 500", "500 to < 1000", "1000 to < 2000", ">= 2000"),
              colors = c("green", "blue", "orange", "red"))
```

## There's lots more
For more details about all the many, many, cool things you can do with the leaflet 
package for R visit http://rstudio.github.io/leaflet/.






