---
title: "Playing with Leaflet"
author: "James Ibbetson"
date: "8/29/2017"
always_allow_html: yes
---

### *Disclaimer*
*The following notes draw heavily from the class lectures and Rmd file covering Leaflet. For the original material and appropriate credit please see the class [Lectures and source code](https://github.com/DataScienceSpecialization/Developing_Data_Products).*

## Introduction

Leaflet is one of the most popular Javascript libraries for
creating interactive maps. The leaflet R package allows you
to create your own leaflet maps without needing to know any
Javascript!

```{r, eval=FALSE}
install.packages("leaflet")
library(leaflet)
```

```{r, echo=FALSE}
library(leaflet)
```

## The basics

Getting started with leaflet is easy. The `leaflet()`
function creates a map widget that you can store in a
variable so that you can modify the map later on. 
```{r, comment = "" }
my_map <- leaflet()
class(my_map)
```

Right now `my_map` is empty. You can add features to the map using 
the pipe operator (`%>%`) just like in dplyr. The `addTiles()` function 
adds mapping data from [Open Street Map](http://www.openstreetmap.org/). 
Add markers to your map using the `addMarkers()` function by 
specifying the longitude and latitude. Add the popup text for when you 
click on the marker with the `popup` argument in `addMarkers()`.

```{r, eval=FALSE}
my_map <- leaflet() %>% 
    addTiles() %>%
    addMarkers(lat=53.430823, lng=-2.960835, popup="This is Anfield")
my_map
```

```{r, echo=FALSE}
my_map <- leaflet() %>% 
    addTiles() %>%
    addMarkers(lat=53.430823, lng=-2.960835, popup="This is Anfield")
my_map
```

## Working with dataframes
The `leaflet()` function is designed to work efficiently with dataframes as inputs.
Multiple markers and their associated information can be passed through 
directly as long as there are columns for latitude and longitude. You can even 
add hyperlinks to the marker popups.
```{r, eval=FALSE}
uk_fun <- data.frame(name = c("Churchill War Rooms", "High Ropes Oxford", "Bletchley Park",
                              "Stonehenge", "Berkhamsted Castle"),
                      lat = c(51.502264, 51.653377, 51.997660, 51.178828, 51.763617),
                      lng = c(-0.129039, -1.232407, -0.740686, -1.826247, -0.559281),
                    sites = c(
  "<a href='http://www.iwm.org.uk/visits/churchill-war-rooms'>Churchill War Rooms</a>",
  "<a href='http://www.highropesoxford.com/'>High Ropes Oxford</a>",
  "<a href='https://www.bletchleypark.org.uk'>Bletchley Park</a>",
  "<a href='http://www.english-heritage.org.uk/visit/places/stonehenge/'>Stonehenge</a>",
  "Berkhamsted Castle"))
uk_fun %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = uk_fun$sites)
```

```{r, echo=FALSE}
uk_fun <- data.frame(name = c("Churchill War Rooms", "High Ropes Oxford", "Bletchley Park",
                              "Stonehenge", "Berkhamsted Castle"),
                      lat = c(51.502264, 51.653377, 51.997660, 51.178828, 51.763617),
                      lng = c(-0.129039, -1.232407, -0.740686, -1.826247, -0.559281),
                    sites = c(
  "<a href='http://www.iwm.org.uk/visits/churchill-war-rooms'>Churchill War Rooms</a>",
  "<a href='http://www.highropesoxford.com/'>High Ropes Oxford</a>",
  "<a href='https://www.bletchleypark.org.uk'>Bletchley Park</a>",
  "<a href='http://www.english-heritage.org.uk/visit/places/stonehenge/'>Stonehenge</a>",
  "Berkhamsted Castle"))
uk_fun %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = uk_fun$sites)
```

## There's lots more
For more details about all the many, many, cool things you can do with the leaflet 
package for R visit http://rstudio.github.io/leaflet/.






