---
title: "My Plotly Notes"
author: "James Ibbetson"
date: "September, 2017"
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### *Disclaimer*
*The following notes draw heavily from the class lectures and Rmd file covering Plotly. For the original material and appropriate credit please see the class [Lectures and source code](https://github.com/DataScienceSpecialization/Developing_Data_Products).*

*Some function notes also come directly from the official R Documentation files.* 

### What is Plotly?
Plotly is a web application for creating and sharing data visualizations. Most importantly, the data plots it creates are interactive.  
  
Plotly can work with several programming languages and applications including R,
Python, and Microsoft Excel.

### Plotly for R
```r
install.packages("plotly")
library(plotly)
```  
An extensive reference for using plotly in R can be found at [https://plot.ly/r/reference/](https://plot.ly/r/reference/).  

The most basic function is `plot_ly()`, with the simple format of `p <- plot_ly(data = df, ~x = x column name, ~y = y column name, type = type of plot)`. However, the exact syntax varies depending on the type of plot so check references. That said, `plot_ly()` will try to create a sensible plot based on the information you give it, inferring as needed.

### A simple scatterplot
```{r, echo=TRUE, message=FALSE}
library(plotly)
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter")
```

### A more complicated scatterplot

By specifying the `color` and `size` arguments you can make each point in your scatterplot a different color and size.

```{r, echo=TRUE, message=FALSE}
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter", 
        color = ~factor(cyl), size = ~hp)
```

### Time series (a plotly aside)
Time series objects are a special dataset type in which multiple observations (i.e. each row of data) are sampled at equispaced points in time. Being such a common and useful type, R has a collection of functions to work with time series.
The function `ts()` is used to create a time series -ts- object from a data vector x. Additional key arguments are start, end, and frequency (respectively, the time of the first observation, last observation, and sampling frequency within the natural time period.)

The function `tsp()` returns the key attributes (in a sense, the meta data) of a ts object.

The functions `start()` and `end()` return the time of the first and last observation. Both return a vector length 2, with the first value being the time period, and the second being the fraction of the time period.

The function `time()` creates a vector of times at which a series was sampled, from the `tsp()` attributes.

The functions `cycle()`, `frequency()`, and `deltat()` return the positions in the cycle of each observation, the number of samples per unit time, and the time interval between observations, respectively.

As a quick example, we can use the `data(airmiles)` data set, which is the revenue passenger miles flown by US airlines for each year from 1937 to 1960:

```{r, echo=TRUE, comment=""}
class(airmiles)
tsp(airmiles)
start(airmiles)
head(airmiles)
head(time(airmiles))
```

Line graphs are the default graph for `plot_ly()`. They're of course useful for
showing change over time:
```{r, echo=TRUE, message=FALSE}
plot_ly(x = time(airmiles), y = airmiles, type = "scatter", mode = "lines")
```

### A list of plotly graph types
The following is a partial list of the `type = "..."` or `type = '...'` argument allowed by `plot_ly()`.<br>
scatter<br>
histogram<br>
box<br>
chloropleth<br>
heatmap<br>
surface<br>
scatter3d<br>
<br>
The last two in the list are 3D plots.<br>
<br>
Also keep in mind that the set of required arguments can change - and be a lot - depending on the `type` selected.

### Multi-line plots
To draw multiple lines on the same graph, plotly requires the data to be a **dataframe** and **organized in long and narrow format**. In other words, all the y data must be in one column, all the x data must be in a second column, and there must be a third column acting like a single factor variable for the set of {x, y}). The latter is passed to the `color = ` argument in `plot_ly()`.

A recommended way to accomplish the required format is by using functions like `gather()` and `mutate()` in the `tidyr` and `dplyr` packages. They are "must use" packages for manipulating data sets. Here's an example:

```{r, echo=TRUE, message=FALSE}
library(plotly)
library(tidyr)
library(dplyr)
data("EuStockMarkets")
stocks <- as.data.frame(EuStockMarkets) %>%
  gather(index, price) %>%
  mutate(time = rep(time(EuStockMarkets), 4))
plot_ly(stocks, x = ~time, y = ~price, color = ~index, type = "scatter", mode = "lines")
```

### Heatmaps
Create a heatmap by using the `type = "heatmap"` argument.

```{r, echo=TRUE, message=FALSE}
terrain1 <- matrix(rnorm(100*100), nrow = 100, ncol = 100)
plot_ly(z = ~terrain1, type = "heatmap")
```


### More plotly resources
- [The Plotly Website](https://plot.ly/)
- [The Plotly R API](https://plot.ly/r/)
- [The Plotly R Package on GitHub](https://github.com/ropensci/plotly)
- [The Plotly R Cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf)
- ["Plotly for R" book by Carson Sievert](https://cpsievert.github.io/plotly_book/)

