---
title: "UK Summer 2017 Sites"
author: "James Ibbetson"
date: "September 2, 2017"
always_allow_html: yes
---

```{r, echo=FALSE}
library(leaflet)
```

## Using leaflet with dataframes
The leaflet package in R is designed to work efficiently with dataframes as inputs for `leaflet()` and its related functions.
Data for multiple markers and their associated information can be passed through 
directly as long as there are columns for latitude and longitude. You can even 
add hyperlinks to the marker popups. Plus many more options...

Here's an example using a simple dataframe with five entries:
```{r, eval=FALSE, warning=FALSE, message=FALSE}
uk_fun <- data.frame(name = c("Churchill War Rooms", "High Ropes Oxford", "Bletchley Park",
                              "Stonehenge", "Berkhamsted Castle"),
                      lat = c(51.502264, 51.653377, 51.997660, 51.178828, 51.763617),
                      lng = c(-0.129039, -1.232407, -0.740686, -1.826247, -0.559281),
                    sites = c(
  "<a href='http://www.iwm.org.uk/visits/churchill-war-rooms'>Churchill War Rooms</a>",
  "<a href='http://www.highropesoxford.com/'>High Ropes Oxford</a>",
  "<a href='https://www.bletchleypark.org.uk'>Bletchley Park</a>",
  "<a href='http://www.english-heritage.org.uk/visit/places/stonehenge/'>Stonehenge</a>",
  "Berkhamsted Castle"))
uk_fun %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = uk_fun$sites) %>%
  setView(lng=-1.2, lat=51.45, zoom = 8)  # specifies map center & zoom
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
uk_fun <- data.frame(name = c("Churchill War Rooms", "High Ropes Oxford", "Bletchley Park",
                              "Stonehenge", "Berkhamsted Castle"),
                      lat = c(51.502264, 51.653377, 51.997660, 51.178828, 51.763617),
                      lng = c(-0.129039, -1.232407, -0.740686, -1.826247, -0.559281),
                    sites = c(
  "<a href='http://www.iwm.org.uk/visits/churchill-war-rooms'>Churchill War Rooms</a>",
  "<a href='http://www.highropesoxford.com/'>High Ropes Oxford</a>",
  "<a href='https://www.bletchleypark.org.uk'>Bletchley Park</a>",
  "<a href='http://www.english-heritage.org.uk/visit/places/stonehenge/'>Stonehenge</a>",
  "Berkhamsted Castle"))
uk_fun %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = uk_fun$sites) %>%
  setView(lng=-1.2, lat=51.45, zoom = 8)  # specifies map center & zoom
```
  
### More
For more details about all the many, many, cool things you can do with the leaflet 
package for R visit http://rstudio.github.io/leaflet/.






