---
title: "Week 3 Assignment (Plotly)"
author: "James Ibbetson"
date: "September 10, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE, comment=""}
knitr::opts_chunk$set(echo = FALSE)
```

## Background

This is an R Markdown presentation demonstrating an interactive plot created with the `plotly` package. It also makes use of the `dplyr` and `tidyr` packages for manipulating dataframes.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(plotly)
library(dplyr)
library(tidyr)
```
<br>
For the example dataset, I chose to download housing price data for California, Florida, North Carolina, and Texas from the [Federal Reserve Bank of St. Louis](https://fred.stlouisfed.org/release?rid=171). The datasets, which cover the period from 1975 to 2017, are available as individual CSV files. 

## Data processing
```{r, echo=TRUE, warning=FALSE, message=FALSE}
allfiles <- lapply(list.files(pattern="*.csv"), 
                   read.csv, stringsAsFactors = FALSE)
df <- as.data.frame(allfiles[1])
for (i in 2:4) df <- inner_join(df, as.data.frame(allfiles[i]))
names(df) <- c("period", "California", "Florida", "NCarolina", "Texas")
head(df)
```

## More data processing
Convert the data to a time series object using `ts()`:
```{r, echo=TRUE, warning=FALSE, message=FALSE}
dts <- ts(df[ , -1], start=c(1975, 1), frequency=4)
```

Convert to the long format dataframe required by `plot_ly()`, and then plot:
```{r, echo=TRUE, eval=FALSE}
prices <- as.data.frame(dts) %>% 
    gather(state, price) %>%
    mutate(year = rep(time(dts), 4))
plot_ly(prices, x = ~year, y = ~price, color = ~state, 
        type = "scatter", mode = "lines")
```

## Relative House Prices in 4 States
1975-2017 (1980 Q1 = 100)
```{r, echo=FALSE}
prices <- as.data.frame(dts) %>% gather(state, price) %>%
    mutate(year = rep(time(dts), 4))
plot_ly(prices, x = ~year, y = ~price, color = ~state, 
        type = "scatter", mode = "lines")
```